

### Objective: Evaluate whether harvest influences presence (or biomass) of a species

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table);
library(mgcv);
library(plotrix)
library(tidyr);
library(Hmisc);
#install.packages('plm');
library(plm);
library(ggplot2);
```





```{r Load datasets}
#setwd("/Users/rebeccaselden/Documents/Collaborations/FishClim")
surv.dat <- readRDS("Data/dat.latbin2.rds")
land.dat <- readRDS("Data/total.land.latbin.rds")
setorder(land.dat, spp, year)
```

### Examine relationship between latitude and SBT
Very strong sigmoid relationship between SBT and latitude
Very strong linear relationship between SST and latitude
Not a surprise that these two variables no longer significant when have spatial fixed effects
--> Clarification from meeting with Malin, the spatial trends in temperature will be soaked up by spatial fixed effects,
but the temporal variation in temperature is what will be evaluated (which is what we specifically want to evaluate for climate change)
```{r temp vs lat}
boxplot(SBT.actual ~ lat.bin, surv.dat[spp=="Clupea harengus"], xlab="Latitude", ylab="SBT.actual")
boxplot(SST.actual ~ lat.bin, surv.dat[spp=="Clupea harengus"], xlab="Latitude", ylab="SST.actual")

```


### Assign catch=0 for areas surveyed and no catch observed 


 Make sure each species has a landing record for each year and lat bin in the survey
 Get rid of southernmost survey bins because not observed in every year

```{r Assign catch to 0}
combined.dat <- merge(surv.dat[!(is.na(lat.bin))& year<2015 &!(lat.bin %in% c("30.5", "31.5", "32.5", "33.5", "34.5"))], 
                      land.dat, by=c("lat.bin", "year", "spp"), all.x=T)
setorder(combined.dat, spp, year, lat.bin)

combined.dat[,"land_mt":=ifelse(is.na(tl_mt), 0, tl_mt)]
```


### Standardize catch so mean of 0 and sd=1 within a species
 If no catch in any year, give value of 0
```{r standardize catch}
catch.scale <- combined.dat[,list(mean_land=mean(land_mt, na.rm=T), sd_land=sd(land_mt, na.rm=T)), by=list(spp)]
combined.dat2 <- merge(combined.dat, catch.scale, by="spp")

combined.dat2[,"land_scaled":=ifelse(mean_land==0, 0, scale(land_mt, center=T, scale=T)), by=list(spp)]

#land_yr <- combined.dat2[,list(mean_land=mean(land_mt), mean_land_scaled=mean(land_scaled)), by=list(spp, year)]
```


### Standardize Temperature within spp

Normalize each temperature for each species range to a mean of 0 and SD of 1 across spp geo range.
 Subtract its mean temperature of occurrence (or of biomass) from the observed temperature.
 Then divide by standard deviation of temperature of occurrence (or of biomass)
```{r examine temperature of occurrence}
par(mfrow=c(2,2))
weighted.hist(combined.dat2[spp=="Gadus morhua"]$SBT.actual, 
              w=combined.dat2[spp=="Gadus morhua"]$wtcpuenal, freq=F, main="Cod SBT, w=Biomass")
weighted.hist(combined.dat2[spp=="Gadus morhua"]$SBT.actual, 
              w=combined.dat2[spp=="Gadus morhua"]$pres2, freq=F, main="Cod SBT, w=Presence")
weighted.hist(combined.dat2[spp=="Centropristis striata"]$SBT.actual, 
              w=combined.dat2[spp=="Centropristis striata"]$wtcpuenal, freq=F, main="Dogfish SBT, w=Biomass")
weighted.hist(combined.dat2[spp=="Centropristis striata"]$SBT.actual, 
              w=combined.dat2[spp=="Centropristis striata"]$pres2, freq=F, main="Dogfish SBT, w=Presence")
```

Calculate mean SBT and SST of presence weighted by presence and biomass
```{r standardize temperature}
combined.dat2[,"mean.SBTpres":=weighted.mean(SBT.actual, w=pres2), by=list(spp)]
combined.dat2[,"mean.SSTpres":=weighted.mean(SST.actual, w=pres2), by=list(spp)]

combined.dat2[,"sd.SBTpres":=sqrt(wtd.var(SBT.actual, weights=pres2)), by=list(spp)]
combined.dat2[,"sd.SSTpres":=sqrt(wtd.var(SST.actual, weights=pres2)), by=list(spp)]

combined.dat2[,"mean.SBTbio":=weighted.mean(SBT.actual, w=wtcpuenal), by=list(spp)]
combined.dat2[,"mean.SSTbio":=weighted.mean(SST.actual, w=wtcpuenal), by=list(spp)]

combined.dat2[,"sd.SBTbio":=sqrt(wtd.var(SBT.actual, weights=wtcpuenal)), by=list(spp)]
combined.dat2[,"sd.SSTbio":=sqrt(wtd.var(SST.actual, weights=wtcpuenal)), by=list(spp)]

combined.dat2[,"scaled.SBT":=(SBT.actual - mean.SBTpres)/sd.SBTpres]
combined.dat2[,"scaled.SST":=(SST.actual - mean.SSTpres)/sd.SSTpres]

plot(wtcpuenal ~ scaled.SBT, combined.dat2[spp=="Gadus morhua"])
```

### Lag variables in time 

Lag catch, biomass, presence, and environment by one, two, and 3 year
```{r lag variables in time}
vars <- c("pres2", "wtcpue", "wtcpuenal", "land_mt", "land_scaled",
          "scaled.SBT", "avg.SBT.min", "avg.SBT.max", 
          "scaled.SST", "avg.SST.min", "avg.SST.max")
lag1cols <- paste("lag1", vars, sep=".")
lag2cols <- paste("lag2", vars, sep=".")
lag3cols <- paste("lag3", vars, sep=".")

lag.dt <- copy(combined.dat2)

lag.dt[,(lag1cols):=shift(.SD, n=1, type="lag"), by=list(spp, lat.bin), .SDcols=vars]
lag.dt[,(lag2cols):=shift(.SD, n=2, type="lag"), by=list(spp, lat.bin), .SDcols=vars]
lag.dt[,(lag3cols):=shift(.SD, n=3, type="lag"), by=list(spp, lat.bin), .SDcols=vars]



### Examine autocorrelation in biomass
plot(wtcpuenal ~ lag1.wtcpuenal, lag.dt[spp=="Gadus morhua"], main="Cod")
abline(a=0, b=1, col="red")

plot(wtcpuenal ~ lag1.wtcpuenal, lag.dt[spp=="Centropristis striata"], main="BSB")
abline(a=0, b=1, col="red")
```






### GAM for Log Bio
Note: depth in a latitude bin does vary by year based on sampling location

Model (log bio) ~ scaled SBT + scaled SST + depth + landings in yr-1 + presence in yr-1 + random(spp)

Test with BSB
```{r gam for logbio}
bsb.mod.bio <- gam(wtcpuenal ~ s(scaled.SBT)+ s(scaled.SST)+ s(depth) + s(lag1.land_scaled) + s(lag1.pres2),
                   family=gaussian, data=lag.dt[spp%in%c("Centropristis striata")])
plot(bsb.mod.bio)
summary(bsb.mod.bio)
```

Linear model with BSB
```{r lm for logbio}
bsb.lm.bio <- lm(wtcpuenal ~ scaled.SBT + scaled.SST + depth + lag1.land_scaled + lag1.pres2,
                 data=lag.dt[spp=="Centropristis striata"])
summary(bsb.lm.bio)
```

```{r Fit biomass GAM models for all species, include=F}
# pdf("Figures/FishClimGAMOut.bio.pdf", height=7, width=5)
# mod.fish.clim.bio <- lag.dt[!(spp %in% catch.scale[mean_land<1]$spp),j={
#   print(spp)
#   t.dt <- .SD
#   num.rec.fished <- length(t.dt$land_mt>0)
#   if(num.rec.fished >5){
#     mod.b <- gam(wtcpuenal ~ s(scaled.SBT)+ s(scaled.SST)+ s(depth) + s(lag1.land_scaled) + s(lag1.pres2),
#                  family=gaussian, data=t.dt)
#     chi <- as.vector(summary(mod.b)$chi.sq)
#     fval <- as.vector(summary(mod.b)$s.table[,3])
#     pval <- as.vector(summary(mod.b)$s.pv)
#     vars <- names(summary(mod.b)$chi.sq)
#     
#     par(mfrow=c(3,2))
#     plot(mod.b)
#     plot(fitted(mod.b), napredict(mod.b$na.action, mod.b$y), main=spp, 
#          ylab="Observed", xlab="Fitted")
#     
#     list(vars=vars, chi=chi, fval=fval, pval=pval)
#   }
#   else{list(vars=NA, chi=NA, fval=NA, pval=NA)}
# }, by=list(spp)]
# dev.off()
```

### Mixed Effects GAM for Log Bio
Species as random effect
See random effects in mgcv https://stat.ethz.ch/R-manual/R-devel/library/mgcv/html/random.effects.html

Model (log bio) ~ scaled SBT + scaled SST + depth + landings in yr-1 + presence in yr-1 + random(spp)

```{r mixed effects GAM logbio}
### All species single model
spplim.mod.bio <- gam(wtcpuenal ~ s(scaled.SBT)+ s(scaled.SST)+ s(depth) + s(lag1.land_scaled) + s(lag1.pres2)+ s(spp, bs="re"),
               family=gaussian, data=lag.dt[spp%in%c("Gadus morhua", "Squalus acanthias", "Limanda ferruginea", "Centropristis striata")])
summary(spplim.mod.bio)
gam.vcomp(spplim.mod.bio)
```

### Beta Regression for Presence

```{r beta regression presence}
### To use beta regression, need pres2 to be (0,1)
lag.dt[,"pres2_beta":=ifelse(pres2==0, 0.00001, ifelse(pres2==1, 0.99999, pres2))]

par(mfrow=c(2,2))
plot(avg.SBT.max ~ lat.bin, lag.dt, ylab="Average SBT Max", xlab="Latitude Bin")
plot(avg.SST.max ~ lat.bin, lag.dt, ylab="Average SST Max", xlab="Latitude Bin")
boxplot(pres2 ~ lat.bin, lag.dt[spp=="Gadus morhua"], main="Atlantic Cod",
        ylab="Mean Presence", xlab="Latitude Bin")
boxplot(pres2 ~ lat.bin, lag.dt[spp=="Limanda ferruginea"], main="Yellowtail Flounder", 
        ylab="Mean Presence", xlab="Latitude Bin")

# par(mfrow=c(3,2))
# plot(pres2 ~ avg.SST.max, lag.dtN[spp=="Gadus morhua"], main="Atlantic Cod",
#      ylab="Mean Presence", xlab="Average SST Max")
# plot(pres2 ~ avg.SST.max, lag.dtN[spp=="Limanda ferruginea"], main="Yellowtail Flounder",
#      ylab="Mean Presence", xlab="Average SST Max")
# plot(pres2 ~ lag.land_scaled, lag.dtN[spp=="Gadus morhua"], main="Atlantic Cod",
#      ylab="Mean Presence", xlab="Scaled Landings t-1")
# plot(pres2 ~ lag.land_scaled, lag.dtN[spp=="Limanda ferruginea"], main="Yellowtail Flounder",
#      ylab="Mean Presence", xlab="Scaled Landings t-1")
# plot(lag.land_scaled ~ lag.pres2, lag.dtN[spp=="Gadus morhua"], main="Atlantic Cod",
#      ylab="Scaled Landings t-1", xlab="Mean Presence t-1")
# plot(lag.land_scaled ~ lag.pres2, lag.dtN[spp=="Limanda ferruginea"], main="Yellowtail Flounder",
#      ylab="Scaled Landings t-1", xlab="Mean Presence t-1")



### Model (pres) ~ max SBT + max SST + depth + landings in yr-1 + presence in yr-1
#### Test with cod
cod.mod <- gam(pres2_beta ~ s(avg.SBT.max)+ s(avg.SST.max)+ s(depth) + s(lag1.land_scaled) + s(lag1.pres2),
               family=betar(link="logit"), data=lag.dt[spp=="Gadus morhua"])

summary(cod.mod)
```

```{r Fit presence GAM for all species, include=F}
# ### Remove unfished and very lightly fished species
# ### use scaled SST & SBT
# pdf("Figures/FishClimGAMOut.pdf", height=7, width=5)
# mod.fish.clim <- lag.dt[!(spp %in% catch.scale[mean_land<1]$spp),j={
#   print(spp)
#   t.dt <- .SD
#   num.rec.fished <- length(t.dt$land_mt>0)
#   if(num.rec.fished >5){
#     mod.p <- gam(pres2_beta ~ s(scaled.SBT)+ s(scaled.SST)+ s(depth) + s(lag1.land_scaled) + s(lag1.pres2),
#                  family=betar(link="logit"), data=t.dt)
#     chi <- as.vector(summary(mod.p)$chi.sq)
#     fval <- as.vector(summary(mod.p)$s.table[,3])
#     pval <- as.vector(summary(mod.p)$s.pv)
#     vars <- names(summary(mod.p)$chi.sq)
#     
#     par(mfrow=c(3,2))
#     plot(mod.p)
#     plot(fitted(mod.p), napredict(mod.p$na.action, mod.p$y), main=spp, 
#          ylab="Observed", xlab="Fitted")
#     
#     list(vars=vars, chi=chi, fval=fval, pval=pval)
#   }
#   else{list(vars=NA, chi=NA, fval=NA, pval=NA)}
# }, by=list(spp)]
# dev.off()
# 
# mod.fish.clim.nopres <- lag.dt[!(spp %in% catch.scale[mean_land<1]$spp),j={
#   print(spp)
#   t.dt <- .SD
#   num.rec.fished <- length(t.dt$land_mt>0)
#   if(num.rec.fished >5){
#     mod.p <- gam(pres2_beta ~ s(avg.SBT.max)+ s(avg.SST.max)+ s(depth) + s(lag1.land_scaled),
#                  family=betar(link="logit"), data=t.dt)
#     chi <- as.vector(summary(mod.p)$chi.sq)
#     pval <- as.vector(summary(mod.p)$s.pv)
#     vars <- names(summary(mod.p)$chi.sq)
#     
#     par(mfrow=c(3,2))
#     plot(mod.p)
#     plot(fitted(mod.p), napredict(mod.p$na.action, mod.p$y), main=spp, 
#          ylab="Observed", xlab="Fitted")
#     qq.gam(mod.p)
#     
#     list(vars=vars, chi=chi, pval=pval)
#   }
#   else{list(vars=NA, chi=NA, pval=NA)}
# }, by=list(spp)]

```


### Regression for Delta Presence

```{r regression delta presence}
lag.dt[,"pres2_delta":=pres2 - lag1.pres2]
lag.dt[,"SBT.delta":=avg.SBT.max - lag1.avg.SBT.max]
lag.dt[,"SST.delta":=avg.SST.max - lag1.avg.SST.max]
lag.dt[,"lag.land.delta":=land_scaled - lag1.land_scaled]

### Low deviance explained for delta temperatures
delta.cod.mod <- gam(pres2_delta ~ s(SBT.delta) + s(SST.delta)+ s(depth) + s(lag1.land_scaled),
                     data=lag.dt[spp=="Gadus morhua"])

### Really low deviance explained for absolute temperatures
delta.cod.mod2 <- gam(pres2_delta ~ s(avg.SBT.max) + s(avg.SST.max)+ s(depth) + s(lag1.land_scaled),
                     data=lag.dt[spp=="Gadus morhua"])
summary(delta.cod.mod2)
```



### First Differences Econometrics Model
Laura found a reference where if not much variation in the explanatory variables over a single time period, can take difference over longer lag (and drop out intervening year?)
If we want to go this route, Laura will read more on it
```{r first differences}
lag.dt[,"bio_delta":=wtcpuenal - lag1.wtcpuenal]
# fd.bio.cod <- gam(bio_delta ~ s(SBT.delta) + s(SST.delta)+ s(lag.land.delta),
#                   data=lag.dt[spp=="Gadus morhua"])
fd.bio.bsb <- lm(bio_delta ~ (SBT.delta) + (SST.delta)+ (lag.land.delta),
                  data=lag.dt[spp=="Centropristis striata"])
summary(fd.bio.bsb)
```
### Manual Fixed Effects Econometrics Model

Substract off mean of each variable for each latitude bin and species

```{r fixed effects}

fe.dt <- lag.dt[,list(m.pres2=mean(pres2, na.rm=T),
                             m.wtcpuenal=mean(wtcpuenal, na.rm=T),
                             m.land =mean(land_scaled, na.rm=T),
                             m.SBT=mean(SBT.actual, na.rm=T),
                             m.SST=mean(SST.actual, na.rm=T)),
                       by=list(spp, lat.bin)]

fe.dt2 <- merge(lag.dt[!(lat.bin %in% c("30.5", "31.5", "32.5", "33.5", "34.5"))],
                fe.dt, by=c("spp", "lat.bin"))

fe.dt2[,"fe.pres2":=pres2 - m.pres2]
fe.dt2[,"fe.wtcpuenal":=wtcpuenal - m.wtcpuenal]
fe.dt2[,"fe.lag.land":=lag1.land_scaled - m.land]
fe.dt2[,"fe.SBT":=SBT.actual - m.SBT]
fe.dt2[,"fe.SST":=SST.actual - m.SST]

fe.mod.bsb <- gam(fe.wtcpuenal ~ s(fe.lag.land) + s(fe.SBT) + s(fe.SST) + s(depth), data=fe.dt2[spp=="Centropristis striata"])
fe.mod.bsb.lm <- lm(fe.wtcpuenal ~ fe.lag.land + fe.SBT + fe.SST + depth, data=fe.dt2[spp=="Centropristis striata"])

summary(fe.mod.bsb)

par(mfrow=c(2,2))
plot(fe.mod.bsb)
# higher landings than average in the previous year associated with higher than average biomass this year 
```


### Panel Data Approaches
 Based on Ethan Addicott's code
 
### Prep the panel data
```{r Panel Data Prep}
library(plm)
df <- combined.dat2[spp=="Centropristis striata"]
setorder(df, lat.bin, year)

### Create the panel
p.df <- pdata.frame(df, index = c("lat.bin","year"))

### Lag landings by one year (in same lat.bin)
p.df$land_mt.lag.1 <- lag(p.df$land_mt, k = 1)

### Lag biomass by one year
p.df$wtcpuenal.lag.1 <- lag(p.df$wtcpuenal, k = 1)

### Lag presence
p.df$lag1.pres2 <- lag(p.df$pres2, k=1)

```

### Compare lags produced by shift vs plm::lag
```{r compare lags}
head(lag.dt[spp=="Centropristis striata" & lat.bin==40.5, 
            list(lag1.land_mt=lag1.land_mt,
                 lag1.wtcpuenal=lag1.wtcpuenal, 
                 SBT.actual=SBT.actual,
                 SST.actual=SST.actual,
                 depth=depth,
                 year=year, lat.bin=lat.bin)])

head(subset(p.df, lat.bin==40.5, 
            select=c("land_mt.lag.1", "wtcpuenal.lag.1", "SBT.actual", "SST.actual", "depth","year")))
```

##### Fixed Effects Panel model 
(similar to what I did manually in the previous section of code). 
 Documentation from plm vignette https://cran.r-project.org/web/packages/plm/vignettes/plm.pdf
The standard way of estimating fixed effects models with, say, group effects entails transforming the data by subtracting the average over time to every variable, which is usually termed time-demeaning. --> model="within"", effect="individiual"
Time effects transforms the data by subtracting the average over the group to every variable= group-demeaning --> model="within", effect="time"
Both would be model="within", effect="twoways"
```{r Panel Fixed Effects for Lat Bin}
m.fe <- plm(wtcpuenal ~ SBT.actual + SST.actual + depth+ wtcpuenal.lag.1+ land_mt.lag.1, data=p.df, model="within", effect="individual")
summary(m.fe)

### Display fixed effects
### Without lagged bio, these were the mean biomass in the lat bin
### With lag.wtcpuenal in the model, these effects are how vary with latitude +/- that predicted from the lagged biomass
fixef(m.fe)

### Test for serial correlation
pbgtest(m.fe) #outcome there still IS serial correlation even with lag.bio in the model!
```

```{r Panel Fixed Effects for Both Time and Individual}
m.fe.both <- plm(wtcpuenal ~ SBT.actual + SST.actual + depth+ land_mt.lag.1, data=p.df, model="within", effect="twoways")
summary(m.fe.both)
```

##### Create instrumental variables for catch to reduce endogeneity with wtcpue
Create position lags
```{r Need to lag position now}
lats <- unique(p.df$lat.bin)
p.df$land_mt.northlag.1 <- NA
for (i in 1:length(lats) - 1) {
for (j in 1972:2014) {
  north.name <- paste(lats[i + 1],"-",j, sep = "")
  name <- paste(lats[i],"-",j, sep = "")
  if (is.na(as.logical(p.df$land_mt[north.name])) == FALSE & is.na(as.logical(p.df$land_mt[name])) == FALSE) {
  p.df$land_mt.northlag.1[name] <- p.df$land_mt[north.name]
  }
  else if (is.na(as.logical(p.df$land_mt[name])) == FALSE)  {
    p.df$land_mt.northlag.1[name] <- NA
  }
}
}


p.df$land_mt.southlag.1 <- NA
for (i in 2:length(lats)) {
for (j in 1963:2014) {
  south.name <- paste(lats[i - 1],"-",j, sep = "")
  name <- paste(lats[i],"-",j, sep = "")
  if (is.na(as.logical(p.df$land_mt[south.name])) == FALSE & is.na(as.logical(p.df$land_mt[name])) == FALSE) {
  p.df$land_mt.southlag.1[name] <- p.df$land_mt[south.name]
  }
  else if (is.na(as.logical(p.df$land_mt[name])) == FALSE)  {
    p.df$land_mt.southlag.1[name] <- NA
  }
}
}
```

Creates time lags for the position lags (eg nlag.ylag.1 is landings in the north, last year)
```{r Time lag position lags}
p.df$land_mt.nlag.ylag.1 <- lag(p.df$land_mt.northlag.1, k = 1)
p.df$land_mt.nlag.ylag.2 <- lag(p.df$land_mt.northlag.1, k = 2)
p.df$land_mt.slag.ylag.1 <- lag(p.df$land_mt.southlag.1, k = 1)
p.df$land_mt.slag.ylag.2 <- lag(p.df$land_mt.southlag.1, k = 2)

```

##### Include instrumental variables in the fixed effects model
 instruments after the |
 
 if the model is y ~x1 + x2 + x3 with x1 and x2 endogenous and z1 and z2 external instruments
 formula=y~x1+x2+x3 | x3 + z1 + z2 Or equivalently
 formula=y~x1+x2+x3 | . -x1-x2+z1+z2 (eg subtract out x1 and x1 and add in z1 and z2)

Instrumental variables with twoway fixed effects 
```{r Instrumental variable with twoway fixed effects}
m.14 <- plm(wtcpuenal~SBT.actual + SST.actual + depth + land_mt.lag.1 | SBT.actual + SST.actual + depth + land_mt.nlag.ylag.1 + land_mt.slag.ylag.1, data = p.df, model = "within", effect = "twoways")
summary(m.14)
```

Instrumental variables with lat bin fixed effect and lagged presence
```{r Instrumental lat bin fe and lagged pres}
m.15 <- plm(wtcpuenal~SBT.actual + SST.actual + depth + lag1.pres2 + land_mt.lag.1 | SBT.actual + SST.actual + depth + land_mt.nlag.ylag.1 + land_mt.slag.ylag.1, data = p.df, model = "within", effect = "individual")
summary(m.15)
```

### Using fixed effects in non-plm framework


To recreate fixed effects but potentially use modeling framework other than linear model, make dummy variable for individual effect that is a factor (rather than lat.bin as continuous variable)
Example: fixed.dum <-lm(y ~ x1 + factor(country) - 1, data=Panel)

Stack exchange about dealing with autocorrelation
https://stats.stackexchange.com/questions/181257/correcting-for-autocorrelation-in-simple-linear-regressions-in-r
Change the model specification to obtain non-autocorrelated errors. For example, run a regression with ARMA errors (easy to implement by arima or auto.arima functions in R including the regressors via the parameter xreg) or -- as DJohnson suggested -- include lags of dependent variable as regressors.
So, is having lagged presence (or biomass) sufficient?
Potentially can evaluate by plotting bio t against bio t-1


Similar (but not identical) results to m.fe
Coefficients for each lat bin very similar to what get with fixef(m.fe)
```{r lat bin dummy}
mod.lbin.dum <- lm(wtcpuenal ~ SBT.actual + SST.actual + depth + lag1.wtcpuenal + lag1.land_mt + lat.bin - 1, data=lag.dt[spp=="Centropristis striata"])

### Using the p.df instead
mod.lbin.dum.p <- lm(wtcpuenal ~ SBT.actual + SST.actual + depth + wtcpuenal.lag.1 + land_mt.lag.1 + lat.bin - 1, data=p.df)



drop1(mod.lbin.dum, test="F")

### Compare with gam
gam.lbin.dum <- gam(wtcpuenal ~ s(SBT.actual) + s(SST.actual) + depth + lag1.wtcpuenal + lag1.land_mt + lat.bin - 1, data=lag.dt[spp=="Centropristis striata"])
```

#### Plot effect using jtools
By default, all numeric predictors other than the one specified in the pred argument are meancentered
```{r plot effect}
mod.cod1 <- lm(wtcpuenal ~ SBT.actual + SST.actual + depth + lag1.land_mt + lag1.land_mt + lat.bin - 1, data=lag.dt[spp=="Gadus morhua" & pres2!=0])

mod.hp1 <- lm(wtcpuenal ~ SBT.actual + SST.actual + depth + lag1.land_mt + lag1.land_mt + lat.bin - 1, data=lag.dt[spp=="Hippoglossoides platessoides" & pres2!=0])

mod.pa1 <- lm(wtcpuenal ~ SBT.actual + SST.actual + depth + lag1.land_mt + lag1.land_mt + lat.bin - 1, data=lag.dt[spp=="Pseudopleuronectes americanus" & pres2!=0])

mod.pm1 <- lm(wtcpuenal ~ SBT.actual + SST.actual + depth + lag1.land_mt + lag1.land_mt + lat.bin - 1, data=lag.dt[spp=="Placopecten magellanicus" & pres2!=0])

mod.lo1 <- lm(wtcpuenal ~ SBT.actual + SST.actual + depth + lag1.land_mt + lag1.land_mt + lat.bin - 1, data=lag.dt[spp=="Leucoraja ocellata" & pres2!=0])

mod.la1 <- lm(wtcpuenal ~ SBT.actual + SST.actual + depth + lag1.land_mt + lag1.land_mt + lat.bin - 1, data=lag.dt[spp=="Lophius americanus" & pres2!=0])

library(jtools)
effect_plot(mod.cod1, pred = lag1.land_mt, interval = TRUE, plot.points = TRUE, main.title="Gadus morhua")
effect_plot(mod.hp1, pred = lag1.land_mt, interval = TRUE, plot.points = TRUE, main.title="Hippoglossoides platessoides")

effect_plot(mod.pm1, pred = lag1.land_mt, interval = TRUE, plot.points = TRUE, main.title="Placopecten magellanicus")
effect_plot(mod.lo1, pred = lag1.land_mt, interval = TRUE, plot.points = TRUE, main.title="Leucoraja ocellata")

effect_plot(mod.la1, pred = lag1.land_mt, interval = TRUE, plot.points = TRUE, main.title="Lophius americanus")



```


### Fit lm with dummy for lat bin all species
```{r lm lat bin dummy all spp}
mod.lbin.dum.all <- lag.dt[,j={
  print(spp)
  t.dt <- .SD
  m <- lm(wtcpuenal ~ SBT.actual + SST.actual + depth + lag1.wtcpuenal + land_mt + lat.bin - 1, data=t.dt)
  list(var=dimnames(summary(m)$coefficients)[[1]],
       est=summary(m)$coefficients[,1],
       pval=summary(m)$coefficients[,4],
       mean_land=unique(mean_land))
}, by=list(spp)]

### 4spp (7 if 0.1) have significant negative coefficient for fishing if use scaled landings 
### 12 spp (19 if 0.1) have positive coefficient
### BSB, summer flounder, and red hake are marginal
### Only 78 species total have a coefficient
### Unfished species don't have a coefficient for scaled landings (because doesn't vary?)
mod.lbin.dum.all[var=="land_mt"&est < 0 & pval<0.05]
mod.lbin.dum.all[var=="land_mt"&est > 0 & pval<0.05]


## 112 of 121 species have lagged bio as significant
mod.lbin.dum.all[var=="lag.wtcpuenal" & pval<0.1]

```

Why Positive effect of fishing on biomass even with lagged biomass in regression
Fish a lot last year, survey catches more than expect from last year's biomass--> why?
If fishing a better index of abundance (because integrated the whole year), maybe + predictor
Maybe lag 2 years for fishing (maybe too much overlap based on when year summed for landings)
or 3 year average? or 3 years separate
Maybe biomass lag 2 years
Or some trend, low catches then bio lower the next year, because both tracking down or up
Wouldn't  lag bio control for the shared trend?
Fish there last year, survey doesn't show (or would be in lagged bio)?
is the fishery catching smaller fish than show up in survey?

### Fit lm with dummy for lat bin all species and scaled landing lagged 2 years for the species
```{r lm lat bin dummy all spp lag2}
mod.lbin.dum.all.lag2 <- lag.dt[,j={
  print(spp)
  m <- lm(wtcpuenal ~ SBT.actual + SST.actual + depth + lag1.wtcpuenal + lag2.land_scaled + lat.bin - 1, data=.SD)
  list(var=dimnames(summary(m)$coefficients)[[1]], 
       est=summary(m)$coefficients[,1],
       pval=summary(m)$coefficients[,4],
       mean_land=unique(mean_land))
}, by=list(spp)]

### 6spp (10 if 0.1) have significant negative coefficient for fishing if use scaled landings 
### 12 spp (14 if 0.1) have positive coefficient
### BSB, summer flounder, and red hake are marginal
### Only 78 species total have a coefficient
### Unfished species don't have a coefficient for scaled landings (because doesn't vary?)
mod.lbin.dum.all.lag2[var=="lag2.land_scaled"&est < 0]
mod.lbin.dum.all.lag2[var=="lag2.land_scaled"&est < 0 & pval<0.1]
mod.lbin.dum.all.lag2[var=="lag2.land_scaled"&est > 0 & pval<0.1]

```


### Fit lm with dummy for lat bin all species and scaled landing lagged 3 years for the species
```{r lm lat bin dummy all spp lag 3}
mod.lbin.dum.all.lag3 <- lag.dt[,j={
  print(spp)
  m <- lm(wtcpuenal ~ SBT.actual + SST.actual + depth + lag1.wtcpuenal + lag3.land_scaled + lat.bin - 1, data=.SD)
  list(var=dimnames(summary(m)$coefficients)[[1]], 
       est=summary(m)$coefficients[,1],
       pval=summary(m)$coefficients[,4],
       mean_land=unique(mean_land))
}, by=list(spp)]

### 6spp (8 if 0.1) have significant negative coefficient for fishing if use scaled landings 
### 11 spp (16 if 0.1) have positive coefficient
### BSB, summer flounder, and red hake are marginal
### Only 78 species total have a coefficient
### Unfished species don't have a coefficient for scaled landings (because doesn't vary?)
mod.lbin.dum.all.lag3[var=="lag3.land_scaled"&est < 0]
mod.lbin.dum.all.lag3[var=="lag3.land_scaled"&est < 0 & pval<0.05]
mod.lbin.dum.all.lag3[var=="lag3.land_scaled"&est > 0 & pval<0.05]

```




The end of the plm vignette shows how many of the fixed effect approaches can be done similarly using a glmm framework (with latitude bin as random effect)
Ben Bolker has a brain dump about how you'd incorporate spatial or temporal correlations in R within glmm
https://bbolker.github.io/mixedmodels-misc/notes/corr_braindump.html

Apparently also possible with gamm
m3 <- gamm(y ~ s(xt, k = 20), correlation = corAR1(form = ~ time))
https://www.fromthebottomoftheheap.net/2011/07/21/smoothing-temporally-correlated-data/
Note: the bad fit with his normal gam was due to the specification of k=20, and the default GCV smoothness selection which doesn't perform as well as ML or REML






