---
title: "PredictCM_RAM"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(raster)
```

### Load Data and restrict to strata observed >90% of years of a survey
```{r set up haul and catch information}
load("Data/hauls_catch_Dec2017.RData", verbose=T)

### Species catch by haul
catch.dt <- as.data.table(dat)
haul.dt <- as.data.table(hauls)

haul.stratum.obs <- haul.dt[,list(num.yrs.obs=length(unique(year))), by=list(stratum, regionfact, ocean, surveyfact)]
yrs.obs.region <- haul.dt[,list(num.yrs.surv=length(unique(year))), by=list(regionfact, ocean, surveyfact)]
haul.stratum.obs2 <- merge(haul.stratum.obs, yrs.obs.region, by=c("regionfact", "ocean", "surveyfact"))
haul.stratum.obs2[,"frac.yrs":=num.yrs.obs/num.yrs.surv]

### Limit to strata observed in >=90% of years of a survey
haul.dt2 <- haul.dt[stratum %in% haul.stratum.obs2[frac.yrs>=0.9]$stratum]
catch.dt2 <- catch.dt[haulid %in% haul.dt2$haulid]

### Merge with haul location
catch.wloc <- merge(catch.dt2, 
                    haul.dt2[,list(haulid=haulid, lat=lat, lon=lon, regionfact=regionfact, year=year, month=month)], 
                    by=c("haulid"))
```

### Predict historical trends based on Jim Morley Clim Fits
Downloaded from Amphiprion nicheMods_PlosOne2018

```{r Predict Historical}
### Fit for all species
# mod_list <- list.files("Data/CEmods", ".RData")
# sppocean_list <- gsub("CEmods_Nov2017_fitallreg_2017_", "", gsub(".RData", "", mod_list))

# pred_list <- vector("list", length(sppocean_list))

### Black Sea Bass only
mod_list <- c("CEmods_Nov2017_fitallreg_2017_centropristis striata_Atl.RData")
sppocean_list <- gsub("CEmods_Nov2017_fitallreg_2017_", "", gsub(".RData", "", mod_list))

pred_list <- vector("list", length(sppocean_list))

for(i in 1:length(mod_list)){
  print(sppocean_list[i])
  load(paste0("Data/CEmods/", mod_list[i]), verbose=T)
  
  par(mfrow=c(2,4))
  plot(mods$mygam1)
  
  dt.fit <- as.data.table(mods$mygam1$model)
  dt.fit.bio <- as.data.table(mods$mygam2$model)
  region.orig <- levels(dt.fit$regionfact) #original fit only had some regions
  
  cols.dt.fit <- colnames(dt.fit)
  haul.dt.lim <- merge(haul.dt2, dt.fit, by=cols.dt.fit[cols.dt.fit != "presfit"])

  pred.fit <- predict(mods$mygam1, newdata=haul.dt.lim, type="response")
  pred.fit.bio <- predict(mods$mygam2, newdata=haul.dt.lim, type="response")
  
  ### Combined prediction (pred.pres * exp(pred.log.bio))
  pred.total<- pred.fit * exp(pred.fit.bio)
  
  pred_list[[i]] <- data.table(haulid=haul.dt.lim$haulid,
                               stratum=haul.dt.lim$stratum,
                               year=haul.dt.lim$year,
                               month=haul.dt.lim$month,
                               regionfact=haul.dt.lim$regionfact,
                               lat=haul.dt.lim$lat,
                               lon=haul.dt.lim$lon,
                               depth=haul.dt.lim$depth,
                               bottemp=haul.dt.lim$bottemp,
                               surftemp=haul.dt.lim$surftemp,
                               SST.seasonal.mean=haul.dt.lim$SST.seasonal.mean,
                               SBT.seasonal=haul.dt.lim$SBT.seasonal,
                               presfit=haul.dt.lim$presfit,
                               pred.pres=pred.fit, 
                               pred.log.bio=pred.fit.bio,
                               pred.total=pred.total,
                               sppocean=sppocean_list[i])
}

pred.dt <- rbindlist(pred_list)
```

### Save output for later work
```{r save output}
#saveRDS(pred.dt, "Output/pred.dt.rds")

#saveRDS(pred.dt, "Output/yellowtail.pred.rds")
```



### Compare prediction and observed for BSB
```{r compare pred obs}
cent.obs <- catch.wloc[,list(lat.cent=weighted.mean(lat, w=wtcpue),
                             lat.cent.pres=weighted.mean(lat, w=presfit)), 
                       by=list(sppocean, regionfact, year)]

cent.cm <- pred.dt[,list(lat.cent.cm =weighted.mean(lat, w=pred.total),
                    lat.cent.pres.cm=weighted.mean(lat, w=pred.pres),
                    sbt=mean(SBT.seasonal, na.rm=T),
                    sst=mean(SST.seasonal.mean, na.rm=T)),
              by=list(sppocean, regionfact, year)]
setorder(cent.cm, year)
```

```{r fig2, fig.height=7, fig.width=7}
par(mfrow=c(2,2), mar=c(4,4,2,2))
plot(lat.cent ~ year, 
     cent.obs[year >=1968 & sppocean=="centropristis striata_Atl" & regionfact=="NEFSC_NEUS"], 
     type="o", main="BSB, NEUS", ylab="Latitude Centroid (w=wtcpue)")
points(lat.cent.cm ~ year,
       cent.cm[year>=1968 & sppocean=="centropristis striata_Atl" & regionfact=="NEFSC_NEUS"],
       type="o", col="blue")
legend("bottomright", legend=c("obs", "pred"), lty=1, col=c("black", "blue"))

### Presence model
plot(lat.cent.pres ~ year, 
     cent.obs[year >=1968 & sppocean=="centropristis striata_Atl" & regionfact=="NEFSC_NEUS"], 
     type="o", main="BSB, NEUS", ylab="Latitude Centroid (w=presence)")
points(lat.cent.pres.cm ~ year,
       cent.cm[year>=1968 & sppocean=="centropristis striata_Atl" & regionfact=="NEFSC_NEUS"],
       type="o", col="blue")
legend("bottomright", legend=c("obs", "pred"), lty=1, col=c("black", "blue"))

plot(sbt ~ year, 
     cent.cm[year >=1968 & year!=2015 & sppocean=="centropristis striata_Atl" & regionfact=="NEFSC_NEUS"], 
     type="o", col="purple", ylim=c(5,14), ylab=c("Temperature"))
points(sst ~ year, cent.cm[year >=1968 & year !=2015 & sppocean=="centropristis striata_Atl" & regionfact=="NEFSC_NEUS"],
       type="o", col="darkgreen", ylim=c(5,14))
legend("bottomright", legend=c("SBT", "SST"), lty=1, col=c("purple", "darkgreen"))
```

BSB is observed to be moving poleward (black), but is not predicted from historical climate conditions to move much (blue)
However, BSB is projected to move poleward substantially based on Ocean Adapt (33->37 degrees latitude across Atlantic)
http://oceanadapt.rutgers.edu/future_data/

Historical SST (green) and SBT (purple) temperatures are increasing






